Invertible Resnets up to Homology Equivalence for Connected Abelian Lie Groups

### 2020-04-04 llncs2e/paper77.tex
errors

\subsection{Deep Learning}
\input{figure2.tex}
Neural nets are used in practice with a different number of perceptrons per layer. However, a layer, from the perspective of a matrix group, can also represent the manifold of the previous one. Manifolds can always be immersed and submersed as long as the rank of the Jacobian of the respective map remains constant. Thus, the dimension of the data manifold is the dimension of the smallest possible bottleneck layer of the neural net  $\dim M = \min_l \dim \textbf{x}^{(l)}$ \citemain{HauserR17}. Our contribution provides the theory of how this dimension can be determined for all neural nets that are a composition of invertible layers acting on an abelian Lie group as domain of the data. The concatenation of such maps is called a deep neural net:
\begin{equation}
    \textbf{x}^{(L)}(M) = \left(\varphi^{(l+1)} \circ \varphi^{(l)} \cdots \circ \varphi^{(1)} \circ \varphi^{(0)} \circ \textbf{x}^{(0)}\right)(M).
\end{equation}
Activation functions restrict to a subgroup of Euclidean space. Continuous homomorphisms between Lie groups are automatically smooth. Thus, a second countable locally Euclidean group has exactly one differentiable structure, turning it into a Lie group. We need to show that $\textbf{x}^{(l+1)}$ has the topology of $G$ and is therefore an embedded closed Lie subgroup $\textbf{x}^{(l)} := H \subseteq G$. Cartans Theorem of Closed Subgroups states that a subgroup $H$ of a Lie group $G$ is an embedded Lie subgroup, if it is a closed subgroup \citemain{cartan1952theorie}. We construct our neural nets such that they are a subset of $\text{GL}(n,\mathbb{R})$ and according to the proof in Appx. \ref{liegroups}, this is a closed subgroup of $\text{GL}(n,\mathbb{C})$, which is itself a Lie group. It follows that a layer in the neural net is also a Lie group. Due to invertibility, the subsequent smaller layers are Lie subgroups, as $\sigma^{(l)}$ is smooth.

@@

\subsection{Results}
Tab. \ref{stats} shows the representatives of homology classes from their respective groups for cifar10 and cifar100. Further, it shows the estimated minimal embedding dimension for a manifold with these homology groups for $\mathcal{C}^1$ to $\mathcal{C}^4$-resnets. The representatives for the $k$-th homology group $\sharp_k^{H}\left(\bar{\mathfrak{L}}\left(\mathfrak{P}(K(X))\right)\right)$ of the respective dataset are counted. Then, we seek for an embedding dimension to represent a manifold with the number of spheres $q$ with these homology groups. A Monte Carlo approach is used to find integer solutions of Eq. \ref{countingequation}. All models below our dimensional treshold show explosions of the gradient, see Fig. \ref{costfunctions}. The dimension of the respective loss function can be identified by the transition \gradient{}, from the lowest tested dimension to the highest, see Fig. \ref{costfunctions}. The models above the treshold remain stable. Higher parameterized tested models are stable. A single net still shows an explosive gradient in Fig. \ref{costfunctions} (b). This could be due to statistical errors in parametrization by mean confidence persistence landscapes. The net showing the deflection of gradient is close to our threshold. We specify the parameterization for an embedding as well as for neural nets of the smoothness degree $1$--$4$. The neural net is capable of eliminating the additional structure of noise, which is no longer part of the abelian Lie group.
\input{figure4.tex}

\section{Conclusions}
\label{sec:conclusion}
Based on the theory of Lie groups and persistent homology a method for invertible resnets was developed to parameterize them with good precision due to the assumption that the data lies on or near by some connected commutative Lie group. This forms an approximate solution for a special case of learning problems, allowing to restrict the domain to a commutative group beeing a connected smooth manifold. The initial problem could be simplified by decomposing the Lie group into a product space $\mathbb{R}^p \times \mathbb{T}^q$. The homology groups of the $q$-torus and the Euclidean space are well known. The application of fully invertible neural nets to resnets demonstrates our results on a different Lie group from Euclidean space. By applying Künneth's theorem the homology groups of the topological factor spaces could be connected with the ones of their product space. Using mean confidence persistence landscapes the homology groups were estimated. A treshold was computed, see Sect. \ref{sec:numericalexperiments}, which is determined by the state space dimension of the resnets according to a representation up to homology equivalence. The mean persistence landscapes are particularly useful for short-lived isolated collections of data points, which in particular generate elements from homology classes at the beginning of the filtration and become visible at the lower left corner of the persistence landscapes, see Fig. \ref{homgroups} (a). The theoretical work was tested with numerical experiments on toy data. We do not limit the processing of the neural net within the layers to an abelian Lie group. Instead, we assume that the net learns some map to a structure close to such an abelian connected Lie group with the estimated homology groups, on which the data could lie. This seems to hold, according to our experiments. Still, this method does only approximate the true data manifold in the case that the data lies already on a product space $\mathbb{R}^p \times \mathbb{T}^q$ or is mapped by the neural network to such a manifold.

>>>>

\subsection{Deep Learning}
\input{figure2.tex}
Neural nets are used in practice with a different number of perceptrons per layer. However, a layer, from the perspective of a matrix group, can also represent the manifold of the previous one. Manifolds can always be immersed and submersed as long as the rank of the Jacobian of the respective map remains constant. Thus, the dimension of the data manifold is the dimension of the smallest possible bottleneck layer of the neural net  $\dim M = \min_l \dim \textbf{x}^{(l)}$ \citemain{HauserR17}. Our contribution provides the theory of how this dimension can be determined for all neural nets acting on an abelian connected Lie group as domain of the data. The concatenation of such maps is called a deep neural net:
\begin{equation}
    \textbf{x}^{(L)}(M) = \left(\varphi^{(l+1)} \circ \varphi^{(l)} \cdots \circ \varphi^{(1)} \circ \varphi^{(0)} \circ \textbf{x}^{(0)}\right)(M).
\end{equation}
Activation functions restrict to a subgroup of Euclidean space. Continuous homomorphisms between Lie groups are automatically smooth. Thus, a second countable locally Euclidean group has exactly one differentiable structure, turning it into a Lie group. We need to show that $\textbf{x}^{(l+1)}$ has the topology of $G$ and is therefore an embedded closed Lie subgroup $\textbf{x}^{(l)} := H \subseteq G$. Cartans Theorem of Closed Subgroups states that a subgroup $H$ of a Lie group $G$ is an embedded Lie subgroup, if it is a closed subgroup \citemain{cartan1952theorie}. We construct our neural nets such that they are a subset of $\text{GL}(n,\mathbb{R})$ and according to the proof in Appx. \ref{liegroups}, this is a closed subgroup of $\text{GL}(n,\mathbb{C})$, which is itself a Lie group. It follows that a layer in the neural net is also a Lie group. Due to invertibility, the subsequent smaller layers are Lie subgroups, as $\sigma^{(l)}$ is smooth.

@@

\subsection{Results}
Tab. \ref{stats} shows the representatives of homology classes from their respective groups for cifar10 and cifar100. Further, it shows the estimated minimal embedding dimension for a manifold with these homology groups for $\mathcal{C}^1$ to $\mathcal{C}^4$-resnets. The representatives for the $k$-th homology group $\sharp_k^{H}\left(\bar{\mathfrak{L}}\left(\mathfrak{P}(K(X))\right)\right)$ of the respective dataset are counted. Then, we seek for an embedding dimension to represent a manifold with the number of spheres $q$ with these homology groups. A Monte Carlo approach is used to find integer solutions of Eq. \ref{countingequation}. All models below our dimensional treshold show explosions of the gradient, see Fig. \ref{costfunctions}. The dimension of the respective loss function can be identified by the transition \gradient{}, from the lowest tested dimension to the highest, see Fig. \ref{costfunctions}. The models above the treshold remain stable. Higher parameterized tested models are stable. A single net still shows an explosive gradient in Fig. \ref{costfunctions} (b). This is due to the fact that the neural network layers themself are not connected nor abelian and are not explicitly restricted to such a structure. The net showing the deflection of gradient is close to our threshold. We specify the parameterization for an embedding as well as for neural nets of the smoothness degree $1$--$4$.
\input{figure4.tex}

\section{Conclusions}
\label{sec:conclusion}
Based on the theory of Lie groups and persistent homology a method for neural networks was developed to parameterize them with good precision due to the assumption that the data lies on or near by some connected commutative Lie group. This forms an approximate solution for a special case of learning problems, allowing to restrict the domain to a commutative group beeing a connected smooth manifold. The initial problem could be simplified by decomposing the Lie group into a product space $\mathbb{R}^p \times \mathbb{T}^q$. The homology groups of the $q$-torus and the Euclidean space are well known. The application of fully invertible neural nets to resnets demonstrates our results on a different Lie group from Euclidean space. By applying Künneth's theorem the homology groups of the topological factor spaces could be connected with the ones of their product space. Using mean confidence persistence landscapes the homology groups were estimated. A treshold was computed, see Sect. \ref{sec:numericalexperiments}, which is determined by the state space dimension of the resnets according to a representation up to homology equivalence. The mean persistence landscapes are particularly useful for short-lived isolated collections of data points, which in particular generate elements from homology classes at the beginning of the filtration and become visible at the lower left corner of the persistence landscapes, see Fig. \ref{homgroups} (a). The theoretical work was tested with numerical experiments on toy data. We do not limit the processing of the neural net within the layers to an abelian Lie group. Instead, we assume that the net learns some map to a structure close to such an abelian connected Lie group with the estimated homology groups, on which the data could lie. This seems to hold, according to our experiments. Still, this method does only approximate the true data manifold in the case that the data lies already on a product space $\mathbb{R}^p \times \mathbb{T}^q$ or is mapped by the neural network to such a manifold.
